A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-10-20 linear.tex
fixed some typos, corrected kappa^2 dependence on dust
corrected kappa^2 dependence on dust
added pressure phase lag
tiny edits

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained. We now discuss explicit solutions to these equations. 
% Numerical solutions are generally required for the 
%vertically-stratified problem.

@@

We solve Eq. \ref{streaming_dispersion} for selected cases where the  
growth rates of the streaming instability have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.

Table \ref{si_compare} compares the eigenfrequencies obtained from the
one-fluid dispersion relation to that from the above studies. As  
expected eigenfrequencies agree better with decreasing $\tstop$ since
in that limit the mixture behaves more like a single fluid. Most 
importantly, we find the work done $\mathcal{W}>0$ in all cases, and
hence find growing oscillations. Thus the streaming instability stems
from the pressure evolution lagging behind that of the total density.

Notice also there is better agreement between eigenfrequencies for
smaller $\mathcal{W}$, i.e. smaller phase pressure-density phase
lags. While this is generally achieved with decreasing $\tstop$, other
problem parameters may affect the magnitude of this phase lag.
For example, `linA' and `linB' have the same $\tstop$ but linA shows
better agreement with the one-fluid growth rate. This suggest that how
well dust and gas are effectively coupled is problem-dependent.

@@

\begin{deluxetable*}{llrrrrr}
  \tablecolumns{7}
  \tablecaption{Selected eigenfrequencies for the streaming
    instability. \label{si_compare}
  }
  \tablehead{
    \colhead{Mode} & 
    \colhead{$\tau_\mathrm{s}\OmK$} &
    \colhead{$\epsilon$} &
     \colhead{$K_x = K_z$} &
      \colhead{$\sigma/\OmK$ (two-fluid)} &
    \colhead{$\sigma/\OmK$ (one-fluid, Eq. \ref{streaming_dispersion})}   &
    \colhead{$\mathcal{W}$ (arbitrary units) } 
  }
\startdata
 linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
 0.4190\ii$ & $0.3647 + 0.4251\ii$ & $0.026$\\

linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
0.0154\ii$&   $-0.4999 + 0.0054\ii$  & $1.74$ \\

linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
0.5981\ii$   &  $0.1344 + 0.6665\ii$  & $3.58\times10^{-4}$\\

linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 +
0.3154\ii$& $0.3228 + 0.3155\ii$ &  $1.84\times10^{-4}$
\enddata
\end{deluxetable*}

>>>>

Eq. \ref{lin_mass}---\ref{lin_energy} is a set of ordinary
differntial equations in $z$. All coefficients and amplitudes are
evaluated at a fiducial radius $r=r_0$, but their full $z$-dependence
is retained. We now discuss solutions to these equations in an unstratified disk 
 for the streaming instability (\S\ref{si}), and a stratified disk for the VSI (\S\ref{results}).

 % Numerical solutions are generally required for the 
%vertically-stratified problem.

@@

{\bf note: kappa2 accounts for dust effect assuming smallh=0.05}
We solve Eq. \ref{streaming_dispersion} for selected cases where the  
growth rates of the streaming instability have been 
calculated analytically from full two-fluid disk models and measured 
from direct numerical particle-gas simulations 
\citep[namely][]{youdin07b,bai10b}. These authors use normalized
wavenumbers $K_x = \eta r k_x$ and similarly for $K_z$, where
\begin{align} 
  \eta \equiv -\frac{1}{2\rhog r\OmK^2}\frac{\p P}{\p r} = 
  \frac{1}{2\left(1-\tepsilon\right)}\frac{F_r}{r\OmK^2}, 
\end{align} 
measures the pressure offset of Keplerian rotation. We use
$\eta=0.05c_s/r\OmK$. They also define the stopping time as 
$\tau_\mathrm{s}=\tstop/(1-\tepsilon)$.

Table \ref{si_compare} compares the eigenfrequencies obtained from the
one-fluid dispersion relation to that from the above studies. As  
expected eigenfrequencies agree better with decreasing $\tstop$ since
in that limit the mixture behaves more like a single fluid. Most 
importantly, we find the work done $\mathcal{W}>0$ in all cases, and
hence find growing oscillations. This implies the streaming instability stems
from the gas pressure evolution lagging behind that of the total (and hence dust) density.

Notice also eigenfrequencies agree better for 
smaller pressure-density phase lags. While this is generally true with decreasing $\tstop$, other 
problem parameters may affect the magnitude of this phase lag, and hence the validity of the one-fluid model.  
For example, `linA' has a larger $\tstop$ than `linB', but linA's one and 
two-fluid growth rates differ by $\sim 1\%$; compare to $\sim 10\%$ error for linB. 
However, linB has a larger phase lag. This means that the effective coupling strength between dust and gas 
is problem-dependent.

@@

\begin{deluxetable*}{llrrrrrr}
  \tablecolumns{8}
  \tablecaption{Selected eigenfrequencies for the streaming
    instability. \label{si_compare}
  }
  \tablehead{
    \colhead{Mode} & 
    \colhead{$\tau_\mathrm{s}\OmK$} &
    \colhead{$\epsilon$} &
     \colhead{$K_{x,z}$} &
      \colhead{$\sigma/\OmK$ (two-fluid)} &
    \colhead{$\sigma/\OmK$ (one-fluid)} &
    \colhead{$\mathcal{W}$ (arbitrary units)} &
      \colhead{Pressure lag}
  }
\startdata
 linA, \cite{youdin07b} &  $0.1$       & 3.0 & 30    & $-0.3480 +
 0.4190\ii$ & $0.3640 + 0.4249\ii$ & $0.026$ & $42\degr$\\

linB, \cite{youdin07b} & $0.1$        &  0.2 & 6 & $0.4999 +
0.0155\ii$&   $-0.4981 + 0.0054\ii$  & $1.75$ & $88\degr$ \\

linC,  \cite{bai10b}  & $10^{-2}$   &  2.0 & 1500&   $0.1049 +
0.5981\ii$   &  $0.1338 + 0.6650\ii$  & $3.57\times10^{-4}$& $78\degr$ \\

linD, \cite{bai10b} &  $10^{-3}$   &  2.0 & 2000 & $0.3225 + 
0.3154\ii$& $0.3219 + 0.3154\ii$ &  $1.85\times10^{-4}$ & $21\degr$ 
\enddata
\end{deluxetable*}
